---
title: "Activity monitoring data analysis"
author: "Yilin"
date: "February 11, 2016"
output: html_document
---
The raw data is collected by a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of Octorber and November,2012 and include the number of steps taken in t minute intervals each day.

Step1: Read the data 
```{r}
rawdata<-read.csv("./activity.csv", sep=",")
```

Step2: histogram showing the total number of steps taken each day
```{r,results="hide"}
sumsteps<-aggregate(steps~date,data=rawdata,sum,na.action=na.omit)
hist(sumsteps[,2],main="Histograme of total steps per day",xlab="Total Steps",fill="blue")
dev.copy(png,width=480,height=480,filename="Histograme of total steps per day")
dev.off()
```

Step3: Mean and median number of steps taken each day
```{r}
meansteps<-aggregate(steps~date,data=rawdata,mean,na.action=na.omit)
mediansteps<-aggregate(steps~date,data=rawdata,median,na.action=na.omit)
head(meansteps)
head(mediansteps)
range(meansteps$steps)
```

Step4: Plot time series plot of the average number of steps taken
```{r,results="hide"}
par(mfrow=c(1,1), mar=c(4,4,2,2))
meansteps$date<-as.Date(meansteps$date)
plot(meansteps,type="o",pch=19,main="Average steps with time",xlab="Date",ylab="Average steps")
dev.copy(png,width=480,height=480,filename="Time series plot of average number of steps taken")
dev.off()
```

Stpe5:Find out the 5-minute interval that, on average,contains the maximum number of steps
```{r}
library(dplyr)
meansteps_interval<-aggregate(steps~interval,data=rawdata,mean,na.action=na.omit)
meansteps_interval<-arrange(meansteps_interval,steps)
max<-meansteps_interval[length(meansteps_interval$steps),]
max
```

Step6:clean the data by substituting "NA"" by the mean of their corresponding interval
```{r}
library(reshape2)
rawdata_t<-melt(rawdata,id=c("interval","date"),na.rm=TRUE)
mean_interval<-dcast(rawdata_t,variable~interval,mean)
mean_interval<-melt(mean_interval,variable.name="interval",value.name="steps")
cleandata<-merge(rawdata,mean_interval[,-1],by="interval",all=TRUE,sort=TRUE)
cleandata$steps.x[is.na(cleandata$steps.x)]<-cleandata$steps.y[is.na(cleandata$steps.x)]
cleandata<-arrange(cleandata[,-4],date,interval,steps.x)
cleandata<-cleandata[,c(3,1,2)]
```

Step7: Plot Histograme of the total number of steps taken each day after missing values are imputed
```{r,results="hide"}
tsteps<-dcast(melt(cleandata,id=c("interval","date")),variable~date,sum)
tsteps<-melt(tsteps,variable.name="date",value.name="total_steps")
hist(tsteps$total_steps,main="Histograme of total steps per day",xlab="Total Steps")
dev.copy(png,width=480,height=480,filename="Histograme of total steps per day_clean data")
dev.off()
```

Step8: Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r,results="hide"}
library(lubridate)
cleandata<-mutate(cleandata,week=weekdays(ymd(date)))
weekend_data<-cleandata[cleandata$week==c("Saturday","Sunday"),]
weekday_data<-cleandata[cleandata$week!=c("Saturday"),]
weekday_data<-weekday_data[weekday_data$week!=c("Sunday"),]
asteps_weekday<-aggregate(steps.x~interval,data=weekday_data,mean)
asteps_weekend<-aggregate(steps.x~interval,data=weekend_data,mean)
par(mfrow=c(1,2),mar=c(4,4,2,2))
plot(asteps_weekday,type="l",main="Week Day",xlab="Interval",ylab="Average steps",ylim=c(0,250))
plot(asteps_weekend,type="l",main="Weekend Day",xlab="Interval",ylab="Average steps",ylim=c(0,250))
dev.copy(png,width=480,height=480,filename="Comparing weekdays and weekends")
dev.off()
```

Summary of the data:

1. Most frequently, the anonymous volunteer takes 1000-1500 steps per day

2. Avarage steps taken by the volunteer at 5 minutes interval per day range from 0.14 to 73.6

3. The maximan steps taken by the volunteer is 206 at the 835th 5 minutes interval

4. After imputing "NA" with the mean steps of its corresponding interval, the data still shows the volunteer most frequently takes 1000-1500 steps per day

5. Comparing the weekdays and weedends, the volunteer takes more steps in weekends. The steps were taken at 500th to 2000th 5 minutes interval.
